---
title: "Hepatisis data analysis"
output:
  pdf_document: default
  html_document:
    number_sections: yes
---

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(visdat)
library(caret)
library(RANN)
library(corrplot)
library(plotly)
library(ggplot2)
library(resample)
library(DataExplorer)
library(imputeMulti)
library(mice)
library(rmarkdown)
library(repr)

library(tidyverse)
library(flextable)
```

# Introduction 

First rows of our data frame.


```{r}
df <- read.table("hepatitis.data", sep = ",")
colnames(df) <- c("Class","Age","Sex","Steroid","Antivirals","Fatigue","Malaise","Anorexia","LiverBig","LiverFirm","SpleenPalpable","Spiders","Ascites","Varices","Bilirubin","AlkPhosphate","Sgot","Albumin","Protime","Histology") # nolint
# attach(df)
```

```{r}
head(df)
```

# Preparing data
```{r}
df[df == "?"] <- NA

df <- mutate_all(df, function(x) as.numeric(as.character(x)))
categorical <- c(1, 3:14, 20)
df[, categorical] <- replace(df[, categorical], df[, categorical] == 2, 0)
df[, categorical] <-  lapply(df[, categorical], as.factor)
```

```{r}
sum(rowSums(is.na(df)) != 0) # we cannot remove these rows!
sum(is.na(df))
vis_dat(df)
# vis_dat(df2)
```

```{r}
plot_missing(df)
```

```{r}
md.pattern(df)
```


```{r}
df.new <- mutate_all(df, function(x) as.numeric(as.character(x)))
data_transform <- preProcess(df.new, method = "knnImpute")
data_transform2 <- preProcess(df.new, method = "bagImpute")
data_transform3 <- preProcess(df.new, method = "medianImpute")

df1 <- predict(data_transform, df.new)
df2 <- predict(data_transform2, df.new)
df3 <- predict(data_transform3, df.new)

df2[c(4:14, 16:17, 19)] <- round(df2[c(4:14, 16:17, 19)])
df3[c(4:14, 16:17, 19)] <- round(df3[c(4:14, 16:17, 19)])

```


```{r echo=FALSE}
for (i in 1:20) {
  column <- df.new[, i]
  if (i %in% c(15, 18)) {
    df1[, i] <- df1[, i] * sd(na.omit(column)) + mean(na.omit(column))
  } else {
    df1[, i] <- round(df1[, i] * sd(na.omit(column)) + mean(na.omit(column)))
  }
}

# categorical <- c(1, 3:14, 20)
# df1[, categorical] <- replace(df1[, categorical], df1[, categorical] == 2, 0)
# df1[, categorical] <-  lapply(df1[, categorical], as.factor)
# df2[, c(2, 16:17, 19)] <- lapply(df2[, c(2, 16:17, 19)], as.integer)
```

```{r results="hide"}
methods = c(" "," "," ","logreg"," ","logreg","logreg","logreg","logreg","logreg","logreg","logreg","logreg","logreg","pmm","pmm","pmm","pmm","pmm"," ")
imp_single <- mice(df, m = 1, method = methods) # Impute missing values
df4 <- complete(imp_single)         # Store imputed data

imp_multi <- mice(df, method = methods)  # Impute missing values multiple times
df5 <- complete(imp_multi, 1)
```

```{r}
# pmm, predictive mean matching (numeric data) 
# logreg, logistic regression imputation (binary data, factor with 2 levels) 
```

```{r echo=FALSE}
df$method <-  c(rep("omit", nrow(df)))
df1$method <- c(rep("knn", nrow(df1)))
df2$method <- c(rep("bag tree", nrow(df2)))
df3$method <- c(rep("median", nrow(df3)))
df5$method <- c(rep("mice", nrow(df5)))
df_all <- rbind(df, df1, df2, df3, df5)
df_all$method <- as.factor(df_all$method)
```
```{r echo=FALSE}
# n <- 9 # 10, 11, 12, 13, 14
# missingness <- c(rep("With Missings", 2), rep("Post Imputation: knnImpute", 2), rep("Post Imputation: bagImpute", 2), rep("Post Imputation: mediumImpute", 2), rep("Post Imputation: mice", 2), rep("Post Imputation: multiply mice", 2)) # Pre/post imputation
# Category <- as.factor(rep(names(table(df[n])), 6))                   # Categories
# Count <- c(as.numeric(table(df[n])), as.numeric(table(df1[n])), as.numeric(table(df2[n])), as.numeric(table(df3[n])), as.numeric(table(df4[n])), as.numeric(table(df5[n])))     # Count of categories
 
# data_barplot <- data.frame(missingness, Category, Count)           # Combine data for plot
 

# ggplot(data_barplot, aes(Category, Count, fill = missingness)) +   # Create plot
#   geom_bar(stat = "identity", position = "dodge") + 
#   scale_fill_brewer(palette = "Set2") +
#   theme(legend.title = element_blank())
```
```{r echo=FALSE}
# ggplot(df_all, aes(x = LiverBig, fill = method)) +   # Create plot
#   scale_fill_brewer(palette = "Set2") +
#   geom_bar(position = "dodge")
```
```{r echo=FALSE}
# ggplot(df_all, aes(x = Protime, fill = method)) +
#   geom_density(alpha = 0.5)
```

```{r fig2, fig.height = 10, fig.width = 10, fig.align = 'center', warning=FALSE, message=FALSE}
ggplot(df_all, aes(x = Protime, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.25) +
  facet_grid(Class ~ .)
```

```{r echo=FALSE}
# col.cat.with.mv <- c(4)
# imputeEM_none <- multinomial_impute(df[col.cat.with.mv], method = "EM", conj_prior = "non.informative", verbose = TRUE)
 ```

# EDA

## Summary
Below we present basic statistics for continuous variables for various types of missing data substitutions. The statics are not significantly different. However, density functions must also be taken into account. They will tell us more about distribution.
```{r echo=FALSE}
stats1 <- df_all %>%
  group_by(method) %>%
  summarize(stat_min = round(min(na.omit(Age)),2),
            stat_q1 = round(quantile(na.omit(Age), 0.25),2),
            stat_median = round(median(na.omit(Age)),2),
            stat_mean = round(mean(na.omit(Age)),2),
            stat_q3 = round(quantile(na.omit(Age), 0.75),2),
            stat_max = round(max(na.omit(Age)),2),
            stat_std = round(sd(na.omit(Age)),2))
stats1$col <- "Age"

stats2 <- df_all %>%
  group_by(method) %>%
  summarize(stat_min = round(min(na.omit(Bilirubin)),2),
            stat_q1 = round(quantile(na.omit(Bilirubin), 0.25),2),
            stat_median = round(median(na.omit(Bilirubin)),2),
            stat_mean = round(mean(na.omit(Bilirubin)),2),
            stat_q3 = round(quantile(na.omit(Bilirubin), 0.75),2),
            stat_max = round(max(na.omit(Bilirubin)),2),
            stat_std = round(sd(na.omit(Bilirubin)),2))
stats2$col <- "Bilirubin"

stats3 <- df_all %>%
  group_by(method) %>%
  summarize(stat_min = round(min(na.omit(AlkPhosphate)),2),
            stat_q1 = round(quantile(na.omit(AlkPhosphate), 0.25),2),
            stat_median = round(median(na.omit(AlkPhosphate)),2),
            stat_mean = round(mean(na.omit(AlkPhosphate)),2),
            stat_q3 = round(quantile(na.omit(AlkPhosphate), 0.75),2),
            stat_max = round(max(na.omit(AlkPhosphate)),2),
            stat_std = round(sd(na.omit(AlkPhosphate)),2))
stats3$col <- "AlkPhosphate"

stats4 <- df_all %>%
  group_by(method) %>%
  summarize(stat_min = round(min(na.omit(Sgot)),2),
            stat_q1 = round(quantile(na.omit(Sgot), 0.25),2),
            stat_median = round(median(na.omit(Sgot)),2),
            stat_mean = round(mean(na.omit(Sgot)),2),
            stat_q3 = round(quantile(na.omit(Sgot), 0.75),2),
            stat_max = round(max(na.omit(Sgot)),2),
            stat_std = round(sd(na.omit(Sgot)),2))
stats4$col <- "Sgot"

stats5 <- df_all %>%
  group_by(method) %>%
  summarize(stat_min = round(min(na.omit(Albumin)),2),
            stat_q1 = round(quantile(na.omit(Albumin), 0.25),2),
            stat_median = round(median(na.omit(Albumin)),2),
            stat_mean = round(mean(na.omit(Albumin)),2),
            stat_q3 = round(quantile(na.omit(Albumin), 0.75),2),
            stat_max = round(max(na.omit(Albumin)),2),
            stat_std = round(sd(na.omit(Albumin)),2))
stats5$col <- "Albumin"

stats6 <- df_all %>%
  group_by(method) %>%
  summarize(stat_min = round(min(na.omit(Protime)),2),
            stat_q1 = round(quantile(na.omit(Protime), 0.25),2),
            stat_median = round(median(na.omit(Protime)),2),
            stat_mean = round(mean(na.omit(Protime)),2),
            stat_q3 = round(quantile(na.omit(Protime), 0.75),2),
            stat_max = round(max(na.omit(Protime)),2),
            stat_std = round(sd(na.omit(Protime)),2))
stats6$col <- "Protime"
all_stats <- bind_rows(stats1, stats2, stats3, stats4, stats5, stats6)
```
```{r echo=FALSE}
numeric_cols <- c("Age", "Bilirubin", "AlkPhosphate", "Sgot", "Albumin", "Protime")
stats <- c("min"," q1", "median", "mean", "q3", "max", "std")

#pivoting
all_stats1 = all_stats[c(1:15),] %>%
    pivot_longer(starts_with("stat")) %>%
    transmute(method, name=paste(col, "-", name), value) %>%
    pivot_wider()

all_stats2 = all_stats[c(16:30),] %>%
    pivot_longer(starts_with("stat")) %>%
    transmute(method, name=paste(col, "-", name), value) %>%
    pivot_wider()

#Creating the table
header_df1 = tibble(col_keys = names(all_stats1),
                   col = c("Method", rep(numeric_cols[1:3],each=7)),
                   stat = c("Method", rep(stats,3)))
header_df2 = tibble(col_keys = names(all_stats2),
                   col = c("Method", rep(numeric_cols[4:6],each=7)),
                   stat = c("Method", rep(stats,3)))

all_stats1 %>% 
    flextable() %>% 
    set_header_df(header_df1) %>%
    merge_v(part="header") %>% 
    merge_h(part="header") %>% 
    theme_box() %>% 
    align(align = "center", part = "header") %>% 
    autofit()

all_stats2 %>% 
    flextable() %>% 
    set_header_df(header_df2) %>%
    merge_v(part="header") %>% 
    merge_h(part="header") %>% 
    theme_box() %>% 
    align(align = "center", part = "header") %>% 
    autofit()
```
## Barplots

```{r}
plot_bar(na.omit(df_all), by = "method", by_position = "dodge")
```

Insights:
* We have more men than women in the dataset. Moreover no women died. This may suggest that men are more likely to die from hepatitis.
* Taking antiviral medications reduces the likelihood of death.
* Almost all of the people who died had symptoms of fatigue and increased liver.
* Histology and spiders might be important (?)
```{r}
df1[, categorical] <-  lapply(df1[, categorical], as.factor)

plot_bar(na.omit(df1[-21]), by = "Class", by_position = "dodge")
```

## Boxplots
```{r}
plot_boxplot(df1, by = "Class")
```
```{r}
numeric_cols <- c("Age", "Bilirubin", "AlkPhosphate", "Sgot", "Albumin", "Protime")

ggplot(df1, aes(x=Albumin, y=Bilirubin,color=Class)) + geom_point(size=3)
```

```{r fig1, fig.height = 10, fig.width = 10, fig.align = 'center'}
cor_matrix <- cor(df2[, sapply(df2, is.numeric)], method = "pearson")
corrplot(cor_matrix, tl.col = "black", addCoef.col = 1, number.cex = 0.9)
```
```{r echo=FALSE}
# plot_histogram(df)
# plot_bar(df)
# plot_qq(df)
# plot_boxplot(df, by = "Class")

## Charts: continuous quantitative data
# Typical questions:
  # 
  # - Uni- or multi-modal distribution?
  # - Symmetric or left- / right-skewed distribution?
  # - Mode / modal interval?
  # - Does the distribution resemble a normal, uniform distribution?

## Charts: discrete quantitative data
# Typical questions:
  #   
  #   - The most frequent value?
  #   - min / max?
  #   - Frequencies of consecutive values?

## Charts: qualitative data
  # Analysis within groups: Why do customers leave?
```
