---
title: "report_addition_Switalska_Kohut.rmd"
output: 
  html_document
date: "2023-12-08"
---
```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(dplyr)
library(caret)
library(mice)
library(DataExplorer)
library(ggpubr)
library(corrplot)
library(imbalance)
library(rpart)
library(rpart.plot)
library(randomForest)
library(ipred)
library(ada)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, results=FALSE}
df <- read.table("hepatitis.data", sep = ",")
colnames(df) <- c("Class","Age","Sex","Steroid","Antivirals","Fatigue","Malaise","Anorexia","LiverBig","LiverFirm","SpleenPalpable","Spiders","Ascites","Varices","Bilirubin","AlkPhosphate","Sgot","Albumin","Protime","Histology")

df[df == "?"] <- NA
df <- mutate_all(df, function(x) as.numeric(as.character(x)))
categorical <- c(1, 3:14, 20)
df[, categorical] <- replace(df[, categorical], df[, categorical] == 2, 0)
df[, categorical] <-  lapply(df[, categorical], as.factor)

df.new <- mutate_all(df, function(x) as.numeric(as.character(x)))
data_transform <- preProcess(df.new, method = "knnImpute")
data_transform2 <- preProcess(df.new, method = "bagImpute")
df1 <- predict(data_transform, df.new)
df2 <- predict(data_transform2, df.new)

unstandarize <- function(data){
  for (i in 1:20) {
    column <- df.new[, i]
    if (i %in% c(15, 18)) {
      data[, i] <- data[, i] * sd(na.omit(column)) + mean(na.omit(column))
    } else {
      data[, i] <- round(data[, i] * sd(na.omit(column)) + mean(na.omit(column)))
    }
  }
  return(data)
}

df2[c(4:14, 16:17, 19)] <- round(df2[c(4:14, 16:17, 19)])
df1 <- unstandarize(df1)

methods = c(" ", " ", " ", "logreg", " ", "logreg", "logreg", "logreg", "logreg", "logreg", "logreg", "logreg", "logreg", "logreg", "pmm", "pmm", "pmm", "pmm", "pmm", " ")
imp_multi <- mice(df, method = methods)  # Impute missing values multiple times
df5 <- complete(imp_multi, 1)

df$method <-  c(rep("omit", nrow(df)))
df1$method <- c(rep("knn", nrow(df1)))
df2$method <- c(rep("bag tree", nrow(df2)))
df5$method <- c(rep("mice", nrow(df5)))
df_all <- rbind(df, df1, df2, df5)
df_all$method <- as.factor(df_all$method)
```

# 3. EDA

## 3.2. Barplots

We can see that bar plots for different methods are very similar. 

```{r, warning=FALSE, message=FALSE}
plot_bar(df_all, by = "method", by_position = "dodge")
```

```{r, warning=FALSE, message=FALSE}
df1[, categorical] <-  lapply(df1[, categorical], as.factor)
plot_bar(df1[-21], by = "Class", by_position = "dodge")
```

```{r, warning=FALSE, message=FALSE}
df2[, categorical] <-  lapply(df2[, categorical], as.factor)
plot_bar(df2[-21], by = "Class", by_position = "dodge")
```

```{r, warning=FALSE, message=FALSE}
df5[, categorical] <-  lapply(df5[, categorical], as.factor)
plot_bar(df5[-21], by = "Class", by_position = "dodge")
```

## 3.3. Histograms

We can see that density plots for different methods are very similar. Only for Protime results are a bit different.

```{r, warning=FALSE, message=FALSE}
p7 <- ggplot(df_all, aes(x = Age, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1)

p8 <- ggplot(df_all, aes(x = Albumin, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1)

p9 <- ggplot(df_all, aes(x = AlkPhosphate, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1)

p10 <- ggplot(df_all, aes(x = Bilirubin, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1)

p11 <- ggplot(df_all, aes(x = Protime, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1)

p12 <- ggplot(df_all, aes(x = Sgot, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1)

ggarrange(p7, p8, p9, p10, p11, p12, ncol = 2, nrow = 3)
```

```{r, warning=FALSE, message=FALSE}
p13 <- ggplot(df_all, aes(x = Age, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1) +
  facet_grid(Class ~ .)

p14 <- ggplot(df_all, aes(x = Albumin, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1) +
  facet_grid(Class ~ .)

p15 <- ggplot(df_all, aes(x = AlkPhosphate, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1) +
  facet_grid(Class ~ .)

p16 <- ggplot(df_all, aes(x = Bilirubin, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1) +
  facet_grid(Class ~ .)

p17 <- ggplot(df_all, aes(x = Protime, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1) +
  facet_grid(Class ~ .)

p18 <- ggplot(df_all, aes(x = Sgot, color = method)) +
  scale_fill_brewer(palette = "Set2") +
  geom_density(size = 1.1) +
  facet_grid(Class ~ .)

ggarrange(p13, p14, p15, p16, p17, p18, ncol = 2, nrow = 3)
```

## 3.4. Q-Q plot

We can see that Q-Q plots for different methods are very similar. Only for Protime results are a bit different.

```{r warning=FALSE, message=FALSE}
plot_qq(df_all, by = "method")
```

## 3.5. Boxplots

We can see that bar plots for different methods are very similar. 

```{r, warning=FALSE, message=FALSE}
plot_boxplot(df_all, by = "method")
```

```{r, warning=FALSE, message=FALSE}
plot_boxplot(df1, by = "Class")
```

```{r, warning=FALSE, message=FALSE}
plot_boxplot(df2, by = "Class")
```

```{r, warning=FALSE, message=FALSE}
plot_boxplot(df5, by = "Class")
```

## 3.6. Correlation
 
We can see that correlation matrices for different methods are very similar. We also added a matrix for initial data with omitting objects with missing values.

```{r, warning=FALSE, message=FALSE}
df_omit <- na.omit(df)
cor_matrix1 <- cor(df_omit[, sapply(df_omit, is.numeric)], method = "pearson")
cor_matrix2 <- cor(df2[, sapply(df2, is.numeric)], method = "pearson")
cor_matrix3 <- cor(df5[, sapply(df5, is.numeric)], method = "pearson")

corrplot(cor_matrix1, tl.col = "black", addCoef.col = 1, number.cex = 0.9)
corrplot(cor_matrix2, tl.col = "black", addCoef.col = 1, number.cex = 0.9)
corrplot(cor_matrix3, tl.col = "black", addCoef.col = 1, number.cex = 0.9)
```


# 4. Classification

```{r, warning=FALSE, message=FALSE, echo=FALSE, results=FALSE}
set.seed(123) 
inTrain <- createDataPartition(y=df1$Class, times=1, p=0.75, list=FALSE)

train1 <- df1[inTrain,-21]
test1 <- df1[-inTrain,-21]
train2 <- df2[inTrain,-21]
test2 <- df2[-inTrain,-21]
train3 <- df5[inTrain,-21]
test3 <- df5[-inTrain,-21]

to_numeric1 <- function(data) {
  data <- mutate_all(data, function(x) as.numeric(as.character(x)))
  return(data)
}

n_new <- sum(train1$Class == 0) - sum(train1$Class == 1)
train.num1 <- to_numeric1(train1)
train.num2 <- to_numeric1(train2)
train.num3 <- to_numeric1(train3)
test.num1 <- to_numeric1(test1)
test.num2 <- to_numeric1(test2)
test.num3 <- to_numeric1(test3)

newMWMOTE1 <- mwmote(train.num1, numInstances = n_new)
train.balanced1 <- rbind(train.num1[-21], newMWMOTE1)
newMWMOTE2 <- mwmote(train.num2, numInstances = n_new)
train.balanced2 <- rbind(train.num2[-21], newMWMOTE2)
newMWMOTE3 <- mwmote(train.num3, numInstances = n_new)
train.balanced3 <- rbind(train.num3[-21], newMWMOTE3)
```

## 4.6. Logistic regression (LR)

```{r, warning=FALSE, message=FALSE, echo=FALSE, results=FALSE}
model.logit1 <-  glm(Class~.-Class, data=train.balanced1, family=binomial(link="logit"))
model.logit2 <-  glm(Class~.-Class, data=train.balanced2, family=binomial(link="logit"))
model.logit3 <-  glm(Class~.-Class, data=train.balanced3, family=binomial(link="logit"))

model.logit4 <-  glm(Class~.-Class-Age-Anorexia-Bilirubin, data=train.balanced1, family=binomial(link="logit"))
model.logit5 <-  glm(Class~.-Class-Age-Anorexia-Bilirubin, data=train.balanced2, family=binomial(link="logit"))
model.logit6 <-  glm(Class~.-Class-Age-Anorexia-Bilirubin, data=train.balanced3, family=binomial(link="logit"))

model.logit7 <-  glm(Class~.-Class-Age-Steroid-AlkPhosphate-Sgot, data=train.balanced1, family=binomial(link="logit"))
model.logit8 <-  glm(Class~.-Class-Age-Steroid-AlkPhosphate-Sgot, data=train.balanced2, family=binomial(link="logit"))
model.logit9 <-  glm(Class~.-Class-Age-Steroid-AlkPhosphate-Sgot, data=train.balanced3, family=binomial(link="logit"))
```

We can see that summary for different methods are similar. 

```{r}
summary(model.logit1)
summary(model.logit2)
summary(model.logit3)

summary(model.logit4)
summary(model.logit5)
summary(model.logit6)

summary(model.logit7)
summary(model.logit8)
summary(model.logit9)
```

## 4.7. Random tree

```{r, echo=FALSE}
mod1 <- Class ~ . - Class
mod2 <- Class ~ . - Class - Age - Steroid - AlkPhosphate - Sgot

to_categorical <- function(data){
  data[, categorical] <-  lapply(data[, categorical], as.factor)
  return(data)
}
train1 <- to_categorical(train1)
train2 <- to_categorical(train2)
train3 <- to_categorical(train3)
train0 <- df[inTrain,-21]
test0 <- df[-inTrain,-21]
train0 <- to_categorical(train0)
test0 <- to_categorical(test0)

set.seed(123) 

full.tree1 <- rpart(mod1, data=train1, control=rpart.control(cp=-1, minsplit=5))
full.tree2 <- rpart(mod1, data=train2, control=rpart.control(cp=-1, minsplit=5))
full.tree3 <- rpart(mod1, data=train3, control=rpart.control(cp=-1, minsplit=5))
full.tree4 <- rpart(mod1, data=train0, control=rpart.control(cp=-1, minsplit=5))
full.tree5 <- rpart(mod2, data=train1, control=rpart.control(cp=-1, minsplit=5))
full.tree6 <- rpart(mod2, data=train2, control=rpart.control(cp=-1, minsplit=5))
full.tree7 <- rpart(mod2, data=train3, control=rpart.control(cp=-1, minsplit=5))
full.tree8 <- rpart(mod2, data=train0, control=rpart.control(cp=-1, minsplit=5))
```

We can see that full trees for different models and different datasets (for different imputation methods) are pretty different.

```{r}
rpart.plot(full.tree1)
rpart.plot(full.tree2)
rpart.plot(full.tree3)
rpart.plot(full.tree4)
rpart.plot(full.tree5)
rpart.plot(full.tree6)
rpart.plot(full.tree7)
rpart.plot(full.tree8)
```

We can see plots and information about misclassification error.

```{r}
plotcp(full.tree1)
printcp(full.tree1)

plotcp(full.tree2)
printcp(full.tree2)

plotcp(full.tree3)
printcp(full.tree3)

plotcp(full.tree4)
printcp(full.tree4)

plotcp(full.tree5)
printcp(full.tree5)

plotcp(full.tree6)
printcp(full.tree6)

plotcp(full.tree7)
printcp(full.tree7)

plotcp(full.tree8)
printcp(full.tree8)
```

```{r, echo=FALSE}
cp.opt1 <- 0.125
cp.opt2 <- 0.083
cp.opt3 <- 0.083
cp.opt4 <- 0.041
cp.opt5 <- 0.083
cp.opt6 <- 0.083
cp.opt7 <- 0.166
cp.opt8 <- 0.062  

full.tree1.pruned <- prune(full.tree1, cp = cp.opt1)
full.tree2.pruned <- prune(full.tree2, cp = cp.opt2)
full.tree3.pruned <- prune(full.tree3, cp = cp.opt3)
full.tree4.pruned <- prune(full.tree4, cp = cp.opt4)
full.tree5.pruned <- prune(full.tree5, cp = cp.opt5)
full.tree6.pruned <- prune(full.tree6, cp = cp.opt6)
full.tree7.pruned <- prune(full.tree7, cp = cp.opt7)
full.tree8.pruned <- prune(full.tree8, cp = cp.opt8)
```

We can see basic information about all pruned trees. There are both similar and different trees.

```{r}
print(full.tree1.pruned)
summary(full.tree1.pruned)

print(full.tree2.pruned)
summary(full.tree2.pruned)

print(full.tree3.pruned)
summary(full.tree3.pruned)

print(full.tree4.pruned)
summary(full.tree4.pruned)

print(full.tree5.pruned)
summary(full.tree5.pruned)

print(full.tree6.pruned)
summary(full.tree6.pruned)

print(full.tree7.pruned)
summary(full.tree7.pruned)

print(full.tree8.pruned)
summary(full.tree8.pruned)
```

## 4.8. Bagging (bootstrap aggregating)

```{r, echo=FALSE}
mod1 <- Class ~ . - Class
mod2 <- Class ~ . - Class - Age - Steroid - AlkPhosphate - Sgot

set.seed(123) 

btree1 <- bagging(mod1, data=train1, nbagg=150, coob=TRUE, minsplit=2, cp=0)
btree2 <- bagging(mod1, data=train2, nbagg=150, coob=TRUE, minsplit=2, cp=0)
btree3 <- bagging(mod1, data=train3, nbagg=150, coob=TRUE, minsplit=2, cp=0)
btree4 <- bagging(mod1, data=train0, nbagg=150, coob=TRUE, minsplit=2, cp=0)

btree5 <- bagging(mod2, data=train1, nbagg=150, coob=TRUE, minsplit=2, cp=0)
btree6 <- bagging(mod2, data=train2, nbagg=150, coob=TRUE, minsplit=2, cp=0)
btree7 <- bagging(mod2, data=train3, nbagg=150, coob=TRUE, minsplit=2, cp=0)
btree8 <- bagging(mod2, data=train0, nbagg=150, coob=TRUE, minsplit=2, cp=0)
```

We can see basic information about all random forests. The error is pretty similar.

```{r}
print(btree1)
print(btree2)
print(btree3)
print(btree4)

print(btree5)
print(btree6)
print(btree7)
print(btree8)
```

## 4.9. Boosting

```{r, echo=FALSE}
mod1 <- Class ~ . - Class
mod2 <- Class ~ . - Class - Age - Steroid - AlkPhosphate - Sgot

set.seed(123) 

boost1 <- ada(mod1, data=train1, iter = 10)
boost2 <- ada(mod1, data=train2, iter = 10)
boost3 <- ada(mod1, data=train3, iter = 10)
boost4 <- ada(mod1, data=train0, iter = 10)
boost5 <- ada(mod2, data=train1, iter = 10)
boost6 <- ada(mod2, data=train2, iter = 10)
boost7 <- ada(mod2, data=train3, iter = 10)
boost8 <- ada(mod2, data=train0, iter = 10)
```

We can see basic information about all random forests. The error is pretty similar.

```{r}
print(boost1)
print(boost2)
print(boost3)
print(boost4)

print(boost5)
print(boost6)
print(boost7)
print(boost8)
```

## 4.10. Random forest

```{r, echo=FALSE}
mod1 <- Class ~ . - Class
mod2 <- Class ~ . - Class - Age - Steroid - AlkPhosphate - Sgot

p1 <-  ncol(test1) - 1
p2 <-  ncol(test1) - 5

set.seed(123) 

rf1 <- randomForest(mod1, data=train1, ntree=500, mtry=p1, importance=TRUE)
rf2 <- randomForest(mod1, data=train2, ntree=500, mtry=p1, importance=TRUE)
rf3 <- randomForest(mod1, data=train3, ntree=500, mtry=p1, importance=TRUE)
rf4 <- randomForest(mod1, data=train1, ntree=500, mtry=sqrt(p1), importance=TRUE)
rf5 <- randomForest(mod1, data=train2, ntree=500, mtry=sqrt(p1), importance=TRUE)
rf6 <- randomForest(mod1, data=train3, ntree=500, mtry=sqrt(p1), importance=TRUE)
rf7 <- randomForest(mod2, data=train1, ntree=500, mtry=p2, importance=TRUE)
rf8 <- randomForest(mod2, data=train2, ntree=500, mtry=p2, importance=TRUE)
rf9 <- randomForest(mod2, data=train3, ntree=500, mtry=p2, importance=TRUE)
rf10 <- randomForest(mod2, data=train1, ntree=500, mtry=sqrt(p2), importance=TRUE)
rf11 <- randomForest(mod2, data=train2, ntree=500, mtry=sqrt(p2), importance=TRUE)
rf12 <- randomForest(mod2, data=train3, ntree=500, mtry=sqrt(p2), importance=TRUE)
```

We can see basic information about all random forests. The error is pretty similar.

```{r}
print(rf1)
print(rf2)
print(rf3)

print(rf4)
print(rf5)
print(rf6)

print(rf7)
print(rf8)
print(rf9)

print(rf10)
print(rf11)
print(rf12)
```

We can see classification error plot for all random forest models. There are both similar and different results

```{r}
plot(rf1)
plot(rf2)
plot(rf3)

plot(rf4)
plot(rf5)
plot(rf6)

plot(rf7)
plot(rf8)
plot(rf9)

plot(rf10)
plot(rf11)
plot(rf12)
```

We can see variable importance ranking for other random forest models. The results are different, but for each model, Protime is in the first place. 

```{r}
# Variable importance ranking
varImpPlot(rf1, main = "Variable Importance Plot")
varImpPlot(rf2, main = "Variable Importance Plot")
varImpPlot(rf3, main = "Variable Importance Plot")

varImpPlot(rf4, main = "Variable Importance Plot")
varImpPlot(rf5, main = "Variable Importance Plot")
varImpPlot(rf6, main = "Variable Importance Plot")

varImpPlot(rf7, main = "Variable Importance Plot")
varImpPlot(rf8, main = "Variable Importance Plot")
varImpPlot(rf9, main = "Variable Importance Plot")

varImpPlot(rf10, main = "Variable Importance Plot")
varImpPlot(rf11, main = "Variable Importance Plot")
varImpPlot(rf12, main = "Variable Importance Plot")
```



